Evolution of pre-copulatory and post-copulatory strategies of inbreeding avoidance

### 2017-08-10 ms_pre_post.Rmd
Introduction edits

In the context of inbreeding avoidance as a key hypothesised driver of reproductive strategy evolution, the absolute and relative frequencies of alleles underlying pre-copulatory and post-copulatory inbreeding avoidance and polyandry will be affected not only by the magnitudes of positive indirect selection stemming from reduced inbreeding depression in females' offspring, but also by the magnitudes of direct negative selection on resulting phenotypes (i.e., the direct fitness costs of expressing each reproductive strategy). Empirical studies have demonstrated diverse costs of mating and mate choice, for example, including energetic costs of developing, maintaining, or enacting necessary physiologies [e.g., @Gasparini2011; @Tuni2013; @Fitzpatrick2014b]; increased risks of predation or disease stemming from increased mate-searching or mating [e.g., @Rowe1988; @Ronkainen1994; @Koga1998]; increased risk of complete mating or fertilisation failure given extreme choosiness [@Kokko2013]; and risks of harm stemming from sexual conflict over fertilisation [e.g., @Rowe1994]. If the relative costs of pre-copulatory and post-copulatory inbreeding avoidance differ, then alleles underlying the less costly strategy might become fixed over generations, while alleles underlying the more costly strategy might go extinct, especially if their effects become redundant following evolution of the less costly strategy. Evolutionary dynamics will therefore depend on the inbreeding avoidance benefit of each strategy, which might also differ and vary in relation to the evolution of the other strategy. Dynamic models that track the frequencies of alleles underlying multiple potentially costly inbreeding avoidance strategies that might influence each other's evolution across generations, and that are enacted among relatives that themselves result from reproductive strategies and inbreeding depression expressed in previous generations, are consequently useful to understand and predict evolutionary outcomes.

We use individual-based modelling to address three key questions regarding evolution of pre-copulatory and post-copulatory inbreeding avoidance and associated polyandry given opportunity for inbreeding and strong inbreeding depression. First, does evolution of post-copulatory inbreeding avoidance alongside pre-copulatory inbreeding avoidance facilitate evolution of polyandry that is otherwise severely constrained by increasing relatedness of polyandrous females to their additional mates? Second, how do costs associated with polyandry and pre-copulatory and post-copulatory inbreeding avoidance affect evolutionary outcomes and, in particular, the long-term persistence of these reproductive strategies given cost asymmetry? Third, how is selection for initially rare pre-copulatory or post-copulatory inbreeding avoidance affected if the other strategy of inbreeding avoidance is already fixed? To address these questions, our model is designed to allow each biological mechanism of interest (i.e., post-copulatory inbreeding avoidance, cost asymmetry, and strategy pre-existence, respectively) to be isolated, thereby facilitating direct comparison of simulations with the mechanism present versus absent with otherwise identical parameter values and conditions. We thereby illustrate how the simultaneous evolution of multiple interacting degenerate reproductive strategies can generate diverse evolutionary outcomes.

>>>>

In the context of inbreeding depression as a key hypothesised driver of reproductive strategy evolution, the absolute and relative frequencies of alleles underlying pre-copulatory and post-copulatory inbreeding avoidance and polyandry will be affected not only by the magnitudes of positive indirect selection stemming from reduced inbreeding depression in females' offspring, but also by the magnitudes of direct negative selection on resulting phenotypes (i.e., the direct fitness costs of expressing each reproductive strategy). Empirical studies have demonstrated diverse costs of mating and mate choice, for example, including energetic costs of developing, maintaining, or enacting necessary physiologies [e.g., @Gasparini2011; @Tuni2013; @Fitzpatrick2014b]; increased risks of predation or disease stemming from increased mate-searching or mating [e.g., @Rowe1988; @Ronkainen1994; @Koga1998]; increased risk of complete mating or fertilisation failure given extreme choosiness [@Kokko2013]; and risks of harm stemming from sexual conflict over fertilisation [e.g., @Rowe1994]. If the relative costs of pre-copulatory and post-copulatory inbreeding avoidance differ, then alleles underlying the less costly strategy might become fixed over generations, while alleles underlying the more costly strategy might go extinct, especially if their effects become redundant following evolution of the less costly strategy. Dynamic models that track the frequencies of alleles underlying multiple, potentially interacting, inbreeding avoidance strategies that are enacted among relatives resulting from reproductive strategies and inbreeding depression expressed in previous generations, are consequently useful to understand and predict evolutionary outcomes.

We use individual-based modelling to address three key questions regarding evolution of pre-copulatory and post-copulatory inbreeding avoidance and associated polyandry given opportunity for inbreeding and strong inbreeding depression. First, does evolution of post-copulatory inbreeding avoidance, alongside pre-copulatory inbreeding avoidance, facilitate evolution of polyandry? Second, how do costs associated with polyandry and pre-copulatory and post-copulatory inbreeding avoidance affect evolutionary outcomes and, in particular, the long-term persistence of these reproductive strategies given cost asymmetry? Third, how is selection for initially rare pre-copulatory or post-copulatory inbreeding avoidance affected if the other strategy of inbreeding avoidance is already fixed? To address these questions, we designed our model to isolate the effect of the biological mechanism of interest (i.e., post-copulatory inbreeding avoidance, cost asymmetry, and strategy pre-existence, respectively) and allow comparison of simulations with the mechanism present versus absent with otherwise identical parameter values and conditions. We thereby illustrate how the simultaneous evolution of multiple interacting degenerate reproductive strategies can generate diverse evolutionary outcomes.
